---
title: "Lab_0"
author: "Matheesha Thambeliyagodage"
date: "February 5, 2017"
output: html_document
---

## On Your Own

In the previous few pages, you recreated some of the displays and preliminary 
analysis of Arbuthnot's baptism data. Your assignment involves repeating these 
steps, but for present day birth records in the United States.  Load up the 
present day data with the following command.

```{r load-present-data, eval=TRUE}
source("C:/Users/MyMats/Documents/Lab0/more/present.R")
present
```

The data are stored in a data frame called `present`.

-   What years are included in this data set? What are the dimensions of the 
    data frame and what are the variable or column names?
    
```{r}
#What years are included in this data set?
present$year

```
```{r}
#What are the dimensions of the data frame
dim(present)

```
```{r}
#what are the variable or column names?
names(present)
```

-   How do these counts compare to Arbuthnot's? Are they on a similar scale?.

```{r}
dim(arbuthnot)
dim(present)

```
No both data sets has similar # of columns but different #of rows


-   Make a plot that displays the boy-to-girl ratio for every year in the data 
    set. What do you see? Does Arbuthnot's observation about boys being born in
    greater proportion than girls hold up in the U.S.? Include the plot in your 
    response.
    
-   When considering below two plots for the "Present" data set for every 100 girls boys both vary betbeen 1046 adn 1058 but in "Arbuthnot's" boys ratio varies between 1050 and 1150.      
```{r}
#boy-to-girl ratio
present$boys/present$girls 
#

```
```{r}
#boy-to-girl ratio Plot for << present >>
plot(main = "Present Boy/Girl Ratio Plot",x = present$year, y = present$boys/present$girls, type = "b")
```
```{r}
#boy-to-girl ratio Plot for << Arbuthnot >>
plot(main = "Arbuthnot's Boy/Girl Ratio Plot", x = arbuthnot$year, y = arbuthnot$boys/arbuthnot$girls, type = "b")
```
```{r}

```

-   In what year did we see the most total number of births in the U.S.?
```{r}
present[present$girls+present$boys == max(present$girls+present$boys), ]$year
```

You can
    refer to the help files or the R reference card 
    [http://cran.r-project.org/doc/contrib/Short-refcard.pdf](http://cran.r-project.org/doc/contrib/Short-refcard.pdf)
    to find helpful commands.

These data come from a report by the Centers for Disease Control 
[http://www.cdc.gov/nchs/data/nvsr/nvsr53/nvsr53_20.pdf](http://www.cdc.gov/nchs/data/nvsr/nvsr53/nvsr53_20.pdf). 
Check it out if you would like to read more about an analysis of sex ratios at 
birth in the United States.

That was a short introduction to R and RStudio, but we will provide you with more
functions and a more complete sense of the language as the course progresses. 
Feel free to browse around the websites for [R](http://www.r-project.org) and 
[RStudio](http://rstudio.org) if you're interested in learning more, or find 
more labs for practice at http://openintro.org.

<div id="license">
This is a product of OpenIntro that is released under a 
[Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel
from a lab written by Mark Hansen of UCLA Statistics.
</div>